---
title: "Task-4 EDA GT"
author: "Shashwat Patel"
date: "06/02/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE,message=FALSE)
```


# **Attaching the packages.**

```{r}
library(tidyverse)
library(gridExtra)
library(plotly)
```

# **Reading the data.**

```{r}
data<-read_csv("globalterrorismdb_0718dist.csv")
```

# **Basic summary of the data**

```{r}
head(data)
tail(data)
```

```{r}
colnames(data)
```

- **Column names of the data which are required.**

- **Extracting the required columns as well**

```{r}
req_col<-c("iyear","country_txt","region_txt","provstate","city","success","suicide","attacktype1_txt","targtype1_txt","natlty1_txt","gname","weaptype1_txt","nkill","nwound")
data<-data[,req_col]
  
```

- Structure of data

```{r}
str(data)
```

- **Removing all rows having NA**

```{r}
data<-data%>%na.exclude()
summary(data)
```

- **Renaming Column names**

```{r}
colnames(data)<-c("Year","Country","Region","State","City","Success","Suicide","Attacktype","Targettype","Nationality","Group","Weapontype","Kills","Wounded")
```


**Exploratory Data Analysis**

- The trend of terrorist attack with year is seen.

- From 1970 to 1992 there was a increase in attack.

- From 1992 to 2003 there was a steady decrease in attack.

- From 2003 to 2014 there was a increase in attack with sudden increase in attack from 2011 to 2014 period.

- From 2014, there is decrease in attack but frequency is still pretty high.

```{r}
pl<-data%>%group_by(Year)%>%summarise(Count=n())%>%ggplot(aes(x=Year,y=Count))+geom_point(colour="blue",size=2)+
  geom_line(colour="red",lwd=1)+ggtitle("Frequency of attack vs Year")+ylab("Frequency")
ggplotly(pl)
```


- Iraq has been greatly affected by terrorist attacks.

- Suprisingly, **United Kingdom** is also featured in this list.



```{r}
top_15_country<-data%>%group_by(Country)%>%summarise(Count=n())%>%slice_max(Count,n=15)
least_15_country<-data%>%group_by(Country)%>%summarise(Count=n())%>%slice_min(Count,n=15)
top_15_country%>%mutate(Country=reorder(Country,Count))%>%ggplot()+geom_col(aes(x=Country,y=Count),fill="#eb4934")+
  coord_flip()+ylab("Frequency")+ggtitle("Top 15 countries affected by terrorist attacks")
```

- Not much can be extracted from this, but this analysis shows that geographical area might play a role. Most of the countries featured here are smaller countries.

```{r}
least_15_country%>%mutate(Country=reorder(Country,Count))%>%ggplot()+geom_col(aes(x=Country,y=Count),fill="#4feb34")+coord_flip()+ggtitle("Least affected 15 countries due to terrorist attack")
```

- Middle East and North African Region is worst affected due to the attacks. South Asia is pretty close also.


```{r}
data%>%ggplot()+geom_bar(aes(x=Region),fill="red")+coord_flip()+ggtitle("Region wise frequency of attack")
```




```{r,fig.height=10,fig.width=10}
data$Success<-factor(data$Success)
data%>%ggplot()+ geom_bar(aes(x=Year,fill=Success),position = "dodge")+ggtitle("Success vs Year")
table(data$Success)
```

- Suicidal attacks are increasing in later years.

```{r,fig.width=10}
data$Suicide<-factor(data$Suicide)
data%>%ggplot()+geom_bar(aes(x=Year,fill=Suicide),position = "dodge",binwidth = 1)+ggtitle("Suicidal attack vs year")
```

- The trend is pretty much similar to Attack count vs year

```{r}
plt2<-data%>%group_by(Year)%>%summarise(Kill=sum(Kills))%>%ggplot(aes(x=Year,y=Kill))+geom_point(colour="blue")+
  geom_line(colour="red")+ggtitle("Kill count vs year" )
ggplotly(plt2)
```

- Majority of attacks is done through Bombing/Explosion or an Armed assault.

```{r}
library(scales)
data%>%group_by(Attacktype)%>%summarise(Kills=sum(Kills))%>%mutate(Attacktype=reorder(Attacktype,Kills))%>%ggplot()+
  geom_col(aes(x=Attacktype,y=Kills),fill="red")+scale_y_continuous(labels = comma)+coord_flip()+ggtitle("Types of attacks and kills due to it")
```

- Majority of target of terrorist are **private citizens and property**

```{r}
data%>%group_by(Targettype)%>%summarise(Kills=sum(Kills))%>%mutate(Targettype=reorder(Targettype,Kills))%>%ggplot()+
  geom_col(aes(x=Targettype,y=Kills),fill="red")+coord_flip()+ggtitle("Kills based on target type")
```



```{r}
data%>%group_by(Nationality)%>%summarise(Kills=sum(Kills))%>%mutate(Nationality=reorder(Nationality,Kills))%>%slice_max(Kills,n=10)%>%ggplot()+geom_col(aes(x=Nationality,y=Kills),fill="#eb4934")+coord_flip()+
  ggtitle("Kills based on Nationality")
```




```{r}
grp_data<-data%>%group_by(Group)%>%summarise(Count=n(),nkills=sum(Kills))
p1<-grp_data%>%mutate(Group=reorder(Group,Count))%>%slice_max(Count,n=7)%>%ggplot()+
  geom_col(aes(x=Group,y=Count),fill="#eb4934")+coord_flip()+ggtitle("Major 7 terrorist groups responsible for attacks")
p2<-grp_data%>%mutate(Group=reorder(Group,nkills))%>%slice_max(nkills,n=7)%>%ggplot()+
  geom_col(aes(x=Group,y=nkills),fill="#eb4934")+coord_flip()+ggtitle("Major 7 groups responsible for killing")
  
grid.arrange(p1,p2,nrow=2)
```

- Majority of people are killed due to explosive and firearms.

```{r}
data%>%group_by(Weapontype)%>%summarise(nkill=sum(Kills))%>%mutate(Weapontype=reorder(Weapontype,nkill))%>%
  slice_max(nkill,n=5)%>%ggplot()+geom_col(aes(x=Weapontype,y=nkill),fill="#eb4934")+coord_flip()+ylab("Kills")+
  ggtitle("Kills based on weapon type")
```



```{r,fig.width=15,fig.height=17}
data%>%ggplot()+geom_bar(aes(Attacktype,fill=Targettype),position = "dodge")
  
```

## Some insight about Country India

```{r}
data_india<-data%>%filter(Country=="India")
```


- From 1970 to 1985 India had very less terrorist attack.

- From 1988 to 1991, it was pretty much constant and then there was a decrease in attack upto 2007

- From 2007, attacks have been increasing steadily at an alarming pace

```{r}
plot<-data_india%>%group_by(Year)%>%summarise(Count=n())%>%ggplot()+geom_point(aes(x=Year,y=Count),colour="blue")+
  geom_line(aes(x=Year,y=Count),colour="red")+ggtitle("Attacks count vs year(India)")
ggplotly(plot)
```

- Srinagar is worst affected city due to attacks.



```{r,fig.width=7}
plt1<-data_india%>%group_by(City)%>%summarise(Count=n())%>%slice_max(Count,n=4)%>%ggplot(aes(x=City,y=Count))+
  geom_col(fill="darkred")+coord_flip()+ggtitle("Worst affetced cities due to attacks")
plt2<-data_india%>%group_by(State)%>%summarise(Count=n())%>%slice_max(Count,n=4)%>%ggplot(aes(x=State,y=Count))+
  geom_col(fill="darkred")+coord_flip()+ggtitle("Worst affected states due to attacks")
grid.arrange(plt1,plt2,nrow=2)
```

# Conclusion

- In most of the 20th Century (1970 to 2000), there was not a significant increase in terrorist attacks, but after 2010 there was seen a significant increase in terrorist attacks.

- **Iraq** is worst affected due to terrorist attacks. Suprisingly, **United Kingdom** is also featured in top 15 worst affected countries due to terrorist attacks. Countries like Iraq, Pakistan are hot spots of terrorism.

- Majority of attacks are in **Middle east, North African and south Asian region**.

- The trend of number of people killed due to terrorist attack in a given year is very similar to frequency of attacks in given year.

- Majority of attacks is done through **bombing/explosion and Armed assault** . The countries should keep a check on illegal selling on explosives as well as on industries dealing with these type of things.

- Majority of targets are **private citizens and property**. It can be inferred that as majority of attacks are done through bombing or explosion, so citizens and property are worst affected. If bombing/explosion is reduced then attack on citzens and property might decrease.

-  Majority of attack and killing is done by unkonwn organization or the other orzanization does not take the responsiblity for the attack. 

- Most of the groups reponsible for higher frequency of attack is also reponsible for killing.

- Number of **suicidal attacks** has increasesd significantly in recent years. There was no trace of suicidal attacks during 1970-2000s.

# Short Conclusion on India

- Worst affected city is Sringar, which is in Jammu and Kashmir. Manipur and Assam are also affected due to the attacks.



